### 2015-01-20 Rebuttal2.txt

1. Why waste time describing LTSes:
We introduced LTSes mainly to define the notion of refinement that we are
using, as well as to define the operators to combine LTSes.

2. The user of a theorem-prover has to walk the "protocol space" through case
analysis:
The theorem-prover has to walk the protocol space through case analysis, but
the proof-script automation in Coq significantly reduces the burden. One can
write proof-scripts that perform a proof-search, i.e. try to apply previously
proven basic lemmas on each of the cases and leave the user with very few
actual proof obligations that are complex or unique so as to not be discharged
using automated proof-search.

3. Invariants are the hardest part in parametric verification of a cache
coherence protocol, and we do not mention it:
We do mention the high level theorems that we use in the paper.  While we
restrict ourselves to the verification of the MSI protocol's implementation,
the proof is modular, and a large portion of the proof can be used, as is, for
verification of other protocols (like MESI, MOSI and MOESI), as we discuss in
the previous rebuttal. The theorem prover helps construct and prove the high
level theorems that make these protocols correct. These theorems depend on a
few basic properties that can be established easily, but were not mentioned in
the paper due to lack of space. These properties include a) state changes occur
when and only when response messages are sent, b) the absence of response
messages from the parent to a child and from the same child to the parent
simultaneously in the network, for the same address and c) FIFO properties of
the responses. These basic properties seem to be the foundation over which the
other cache coherence protocols are built -- they eliminate the corner cases
simplifying the proof. For example, since responses from a child to its parent
and from the parent to the same child for the same address cannot be present
simultaneously, an upgrade response from the parent cannot cross a downgrade
response from the child, which would have eventually led to the upgrade of the
child and downgrade of the parent violating Lemma 3. As pointed out by
Reviewer-E, the theorems and these properties serve as a method to derive a
correct protocol (the different protocols are different system instances which
obey these properties).

4. Pipelining protocol engines and Bluespec:
Though the semantics of the atomic transitions in our paper execute one at a
time, different independent transitions can be made to occur in parallel in the
actual implementation without changing the semantics; Bluespec compiler excels
at this optimization.  One can take this further, where each atomic transition
can be spread across multiple hardware clock cycles, while ensuring that the
overall atomic transaction semantics are obeyed. Bluespec implementation
currently does not include this optimization, though detailed algorithms to
accomplish this have been published have been published [Karczmarek, MEMOCODE'14].

@@

5. Nirav's point: I don't see that Reviewer-E's scathing concern fall
into the core of what we're addressing; our focus on modularity is to
allow verification to be (mostly) carried over to different
microarchitectural variants, which is a bare minimum if one wants
architects to have any lateral freedom to design a working
implementation with a low-cost path to verification. I don't see why
there's any fundamental reason that "scalable" approaches he's pointed to
can't be fit into our methodology directly? I've only glanced at the
papers, but it seems like it should be so.
--- Note sure how to condense it and incorporate -Murali

Reviewer-F:
1. Evidence of proofs being machine-checked:
We have submitted the source code of our Coq development as a supplement to our
paper in the submission website.

2. Figure 3, Tags in Figures 5 & 6:
Both these points are valid and we should have indicated them explicitly.

@@

@@@ Murali response
A new non-conventional cache coherence protocol that produces sequentially
consistent results for data-rate free programs. There's no formal proof that
this protocol is correct...
(Needs to be read further, but doesn't seem relevant)
***

The authors claim that they are achieving their proof for an arbitrary number
of processors. They do not mention anything about the invariants involved in
their parametric proofs. This leads me to suspect that these invariants are
pretty straightforward.  If one harks back to work addressing real protocols,
the Chou, Mannava, Park work
(http://scholar.google.com/citations?view_op=view_citation&hl=en&user=ufGYoL0AAAAJ&citation_for_view=ufGYoL0AAAAJ:qjMakFHDy7sC)
shows how exceedingly hard it is to discover these invariants. In fact, the CMP
approach is an incremental CEGAR-style method for discovering these invariants.
Stating these invariants explicitly would be virtually impossible - even for
FLASH-like protocols.  In fact, a whole series of papers have tried to address
invariant discovery in this space : http://dl.acm.org/citation.cfm?id=1517434

@@@ Murali response
Chou, Mannava, Park [CMP-method] approach uses three paradigms: a) symmetry of
cache coherence protocols b) abstraction of several nodes which behaves as just
another node and c) counter-example guided abtraction refinement (CEGAR) for
refining the abstraction's properties. Using CEGAR, they get invariants for the
rest of the system while concentrating on a finite number of nodes (2).  Each
refinement will not restrict the behavior, but will help towards proving a
safety property. Specifically, let P be an invariant for the abstract
transition system, and we should prove that it satisfies safety property X. We
start out with a P that doesn't satisfy X. Then we add constraints on P, while
permitting the same transitions. Eventually P will satisfy X. The manual
portion of this procedure is in strengthening P.
***

@@@ Murali response
The second paper tries to formulate better invariants in the first place using
the notion of "flows". It says what other messages can interfere when the other
message is being processed...
(Needs to be read further -- I don't understand "flows" clearly; can't say
anything more before I understand that properly)
***

### 2015-01-20 Rebuttal2.txt

1. Why waste time describing LTSes:
We introduced LTSes mainly to define the notion of refinement that we are
using, as well as to define the operators to combine LTSes.

adamc suggests: This description here sounds like the wrong level of detail.  If we respond to the point, IMO it should be just to say that we wanted to review standard material for the benefit of readers unfamiliar with it; the point wasn't to present novel concepts.

@@

2. The user of a theorem-prover has to walk the "protocol space" through case
analysis:
The theorem-prover has to walk the protocol space through case analysis, but
the proof-script automation in Coq significantly reduces the burden. One can
write proof-scripts that perform a proof-search, i.e. try to apply previously
proven basic lemmas on each of the cases and leave the user with very few
actual proof obligations that are complex or unique so as to not be discharged
using automated proof-search.

3. Invariants are the hardest part in parametric verification of a cache
coherence protocol, and we do not mention it:
We do mention the high level theorems that we use in the paper.  While we
restrict ourselves to the verification of the MSI protocol's implementation,
the proof is modular, and a large portion of the proof can be used, as is, for
verification of other protocols (like MESI, MOSI and MOESI), as we discuss in
the previous rebuttal. The theorem prover helps construct and prove the high
level theorems that make these protocols correct. These theorems depend on a
few basic properties that can be established easily, but were not mentioned in
the paper due to lack of space. These properties include a) state changes occur
when and only when response messages are sent, b) the absence of response
messages from the parent to a child and from the same child to the parent
simultaneously in the network, for the same address and c) FIFO properties of
the responses. These basic properties seem to be the foundation over which the
other cache coherence protocols are built -- they eliminate the corner cases,
simplifying the proof. For example, since responses from a child to its parent
and from the parent to the same child for the same address cannot be present
simultaneously, an upgrade response from the parent cannot cross a downgrade
response from the child, which would have eventually led to the upgrade of the
child and downgrade of the parent, violating Lemma 3. As pointed out by
Reviewer-E, the theorems and these properties serve as a method to derive a
correct protocol (the different protocols are different system instances which
obey these properties).

4. Pipelining protocol engines and Bluespec:
Though the semantics of the atomic transitions in our paper execute one at a
time, different independent transitions can be made to occur in parallel in the
actual implementation without changing the semantics; Bluespec compiler excels
at this optimization.  One can take this further, where each atomic transition
can be spread across multiple hardware clock cycles, while ensuring that the
overall atomic transaction semantics are obeyed. Bluespec implementation
currently does not include this optimization, though detailed algorithms to
accomplish this have been published [Karczmarek, MEMOCODE'14].

@@

3. Figure 6: The third case:
Reviewer-F's assertion is not true. The "upd" function merely determines if the
"real" state of the system is as dictated by the state (s,pc) or if it should
be modified to take into account the responses from the memory system. If the
processor is not waiting for the memory system response, then the state is
exactly (s,pc). We made an optimization in the fourth case where we make the
real state H whenever we know that the next instruction we decode is H; instead
the state can be left as (s,pc) in the fourth case.

4. Figure 8: Tags for non-speculative loads:
Our invariant is that there can be only one outstanding non-speculative load,
though there can be several outstanding speculative loads along with the
outstanding non-speculative load. Thus, anything without a tag is the
non-speculative load and there is only one of them, and anything with a tag is
a speculative load.

Reviewer-E's related work summary:

FM methodology: The methodology used for proving that the memory system
realizes cache coherence is not new. A theorem-proving approach to verifying
memory systems has been tacit in many publications. Researchers preferred model
checking, as opposed to pushing through theorem-proving based proofs, largely
because of the daunting complexity of realistic memory system protocols. One
can find several approaches that address specific aspects of memory system
verification in, say, Wim Hesselink's work :
http://www.informatik.uni-trier.de/~ley/pers/hd/h/Hesselink:Wim_H=.

### 2015-01-20 Rebuttal2.txt

Reviewer-E

1. On Describing LTSes:

We include the LTS description for the benefit of readers unfamiliar with it;
the point wasn't to present novel concepts.

2. Modularity

 One of the major benefits of our approach is that it unifies the
 modularization of the design and its verification task which enables
 verification to be done in tandem with mircoarchitectural design as noted by
 Reviewer-E.  Further, this allows us to require only incremental verification
 effort for an incremental microarchitectural change, enabling scaling of the
 verification task.

3. Model Checking vs. Theorem Proving

a. Finding invariants is fundamentally hard. In [PAPER],

Finding invariants is fundamentally hard. We are equivalent in complexity but
  because of our modularization we have a free source of abstraction from the
  designer.

   - In comparision, Quote paper. Invariants were dervied manually, took one
  month, and relied on a manual abstraction of the system space based on
  symettry. Our work is similar save that we do not require an abstraction.

  In comparision to the <PAPER>, our invariants of similar complexity. In
  general, we believe that such invariants relatively simple but subtle and
  generally intuitively understood by the designers.

b. On Case-Analysis:

Much of the requisite protocol case analysis is done automatically via
proof-script automation in Coq. In our experience, cases where proof-script
automation fail required an additional invariant; we strongly believe that
manually providing a similiar invariant are necessary in a model checking
approach.

c. Benefits of TP-based parameterization: It should also be noted that their
  work while hierarchical is only 2-level while ours applies to arbitrary
  levels of hierarchy.

4. Pipelining protocol engines and Bluespec:

While atomic transitions in our paper have one-at-a-time semantics, the
Bluespec compiler generates hardware which executes independant transitions
concurrently in a clock cycle. The Bluespec compiler excels at this
optimization. Atomic transition can also be spread across multiple clock
cycles, while maintaining atomic transaction semantics [Karczmarek,
MEMOCODE'14].

Reviewer-F:

1. Evidence of proofs being machine-checked:

We submitted the source code of our Coq development as a supplement to our
paper in the submission website.

2. Figure 3, tags in Figures 5 & 6:
Both these points are valid and we should have indicated them explicitly.

3. Figure 6: Clarification of the third case: The "upd" function merely
determines if the "real" state of the system is as dictated by (s,pc) or if it
should be modified to take into account memory responses. If the processor
is not waiting for a memory response, then the state is exactly (s,pc). We made
an optimization in the fourth case where we make the real state H whenever we
know that the next instruction we decode is H; we could have it as (s,pc)
instead.

4. Figure 8: Tags for non-speculative loads: Our invariant is that there can be
only one outstanding non-speculative load, though there can be several
outstanding speculative loads concurrently. Anything without a tag is the
current non-speculative load, and anything with a tag is a speculative load.

========================

4. Invariants are the hardest part in parametric verification of a cache
coherence protocol, and we do not mention it:
We do mention the high level theorems that we use in the paper.  While we
restrict ourselves to the verification of the MSI protocol's implementation,
the proof is modular, and a large portion of the proof can be used, as is, for
verification of other protocols (like MESI, MOSI and MOESI), as we discuss in
the previous rebuttal. The theorem prover helps construct and prove the high
level theorems that make these protocols correct. These theorems depend on a
few basic properties that can be established easily, but were not mentioned in
the paper due to lack of space. These properties include a) state changes occur
when and only when response messages are sent, b) the absence of response
messages from the parent to a child and from the same child to the parent
simultaneously in the network, for the same address and c) FIFO properties of
the responses. These basic properties seem to be the foundation over which the
other cache coherence protocols are built -- they eliminate the corner cases,
simplifying the proof. For example, since responses from a child to its parent
and from the parent to the same child for the same address cannot be present
simultaneously, an upgrade response from the parent cannot cross a downgrade
response from the child, which would have eventually led to the upgrade of the
child and downgrade of the parent, violating Lemma 3. As pointed out by
Reviewer-E, the theorems and these properties serve as a method to derive a
correct protocol (the different protocols are different system instances which
obey these properties).

=====================================================================================================
= Our Notes
=====================================================================================================

### 2015-01-21 Rebuttal2.txt

One of the major benefits of our approach is that it unifies the modularization
of the design and its verification task which enables verification to be done
in tandem with mirco-architectural design (as noted by Reviewer-E). Further,
this allows us to require only incremental verification effort for an
incremental microarchitectural change, enabling scaling of the verification
task.

Our approach vs. Model-Checking-based:

a) Deriving Invariants:
Finding invariants is fundamentally hard. The CMP-method (referred by
Reviewer-E) uses symmetry to abstract the system, and then inserts invariants
(called non-interference-lemmas) manually to eliminate counter-examples that
violate the required property (the flow-based methodology [1] specifies these
non-interference-lemmas in the form of properties of action sequences that
takes place for any high-level transaction.) All these invariants are
relatively simple but very subtle and are well understood by the designers. For
example, for the FLASH protocol, one of the non-interference-lemma (Lemma-5 in
[2]) says that whenever a cache is in E state, then the data it has is
up-to-date. This is very similar to our Lemma-1. In general, we believe that
the level of complexity of the lemmas in the two approaches are similar.

The verification of the hierarchical protocol in [3] is restricted to 2-levels
whereas our work applies to arbitrary levels of hierarchy. We are not sure if
it's a fundamental restriction on the CMP-method, since the kind of symmetry
that CMP-method exploits in a one or two-level system is not present in a
hierarchical system.

b) Case-Analysis:
Much of the requisite protocol case analysis is done automatically via
proof-script automation in Coq. In our experience, cases where proof-script
automation fail required an additional invariant; we believe that manually
providing a similiar invariant is necessary even in a model-checking approach.

@@

Reviewer-F:

Figure-3, no tags in Figures-5,6:
Both these points are valid and we should have indicated them explicitly.

Figure-6, third case:
The "upd" function merely determines if the "real" state of the system is as
dictated by (s,pc) or if it should be modified to take into account memory
responses. If the processor is not waiting for a memory response, then the
state is exactly (s,pc). We made an optimization in the fourth case where we
make the real state H whenever we know that the next instruction we decode is
H; we could have it as (s,pc) instead.

Figure-8: tags for non-speculative loads:
Our invariant is that there can be only one outstanding non-speculative load,
though there can be several outstanding speculative loads concurrently.
Anything without a tag is the current non-speculative load, and anything with a
tag is a speculative load.

Reviewer-E:

Pipelining protocol-engines:
While atomic transitions in our paper have one-at-a-time semantics, the
Bluespec compiler generates hardware which executes independant transitions
concurrently in a clock cycle (as the semantics aren't violated); the compiler
excels at this optimization. Atomic transition can also be spread across
multiple clock cycles, while maintaining the semantics [4].

Evidence of machine-checked proofs:
We submitted our Coq source code as a supplement to our paper in the submission
website.

On Describing LTSes:
We include the LTS description for the benefit of readers unfamiliar with it;
the point wasn't to present novel concepts.

References:
[1] Parameterized Verification using Message Flows: An Industrial Experience,
John O'Leary, Murali Talupur, Mark R. Tuttle, FMCAD'09
[2] A Simple Method for Parameterized Verification of Cache Coherence Protocols,
Ching-Tsun Chou, Phanindra K. Mannava, Seungjoon Park, FMCAD'04
[3] Efficient methods for formally verifying safety properties of hierarchical
cache coherence protocol, Xiaofang Chen, Yu Yang, Ganesh Gopalakrishnan,
Ching-Tsun Chou, FMSD'10
[4] A new synthesis procedure for atomic rules containing multi-cycle function
blocks, Michal Karczmarek, Arvind, Muralidaran Vijayaraghavan, MEMOCODE'14

### 2015-01-21 Rebuttal2.txt

One of the major benefits of our approach is that it unifies the modularization
of the design and its verification task which enables verification to be done
in tandem with mirco-architectural design (as noted by Reviewer-E). Further,
this allows us to require only incremental verification effort for an
incremental microarchitectural change, enabling scaling of the verification
task.

Our approach vs. Model-Checking-based:

a) Deriving Invariants:
Finding invariants is fundamentally hard. The CMP-method (referred by
Reviewer-E) uses symmetry to abstract the system, and then inserts invariants
(called non-interference-lemmas) manually to eliminate counter-examples that
violate the required property (the flow-based methodology [1] specifies these
non-interference-lemmas in the form of properties of traces that
takes place for any high-level transaction.) All these invariants are
relatively simple but very subtle and are well understood by the designers. For
example, for the FLASH protocol, one of the non-interference-lemma (Lemma-5 in
[2]) says that whenever a cache is in E state, then the data it has is
up-to-date. This is very similar to our Lemma-1. In general, we believe that
the level of complexity of the lemmas in the two approaches are similar.

@@

The verification of the hierarchical protocol in [3] is restricted to 2-levels
whereas our work applies to arbitrary levels of hierarchy. We are not sure if
it's a fundamental restriction on the CMP-method, since the kind of symmetry
that CMP-method exploits in a one or two-level system is not present in a
hierarchical system.

b) Case-Analysis:
Much of the requisite case-analysis is done automatically via proof-script
automation in Coq. In our experience, cases where proof-script automation fail
required an additional invariant; we believe that manually providing a similiar
invariant is necessary even in a model-checking approach.

### 2015-01-21 Rebuttal2.txt

Reviewer-E:

Pipelining protocol-engines:
While atomic transitions in our paper have one-at-a-time semantics, the
Bluespec compiler generates hardware which executes independant transitions
concurrently in a clock cycle (as the semantics aren't violated); the compiler
excels at this optimization. Atomic transition can also be spread across
multiple clock cycles, while maintaining the semantics [4].

Evidence of machine-checked proofs:
We submitted our Coq source code as a supplement to our paper in the submission
website.

Beyond 500 words:
On Describing LTSes:
We include the LTS description for the benefit of readers unfamiliar with it;
the point wasn't to present novel concepts.

### 2015-01-21 Rebuttal2.txt

One of the major benefits of our approach is that it unifies the modularization
of the design and its verification task which enables verification to be done
in tandem with mirco-architectural design (as noted by Reviewer-E). Further,
this allows us to require only incremental verification effort for an
incremental microarchitectural change, enabling scaling of the verification
task.

Our approach vs. Model-Checking-based:

a) Deriving Invariants:
Finding invariants is fundamentally hard. The CMP-method referred by Reviewer-E
uses symmetry to abstract the system, and then inserts invariants, called
non-interference-lemmas, manually to eliminate counter-examples that violate
the required property (the flow-based methodology [1] specifies invariants in
the form of properties of traces that take place for any high-level
transaction.) All these invariants are relatively simple but very subtle in
terms of their efficacy. For example, for the FLASH protocol, one of the
non-interference-lemma (Lemma-5 in [2]) says that whenever a cache is in E
state, then the data it has is up-to-date. This is very similar to our Lemma-1.
In general, we believe that the level of complexity of the lemmas in the two
approaches are similar. The invariants are intuitively understood at the time
of design of each module, making the task of formalizing them during the
module's verification manageable.

The verification of the hierarchical protocol in [3] is restricted to 2-levels
whereas our work applies to arbitrary levels of hierarchy. We are not sure if
it's a fundamental restriction on the CMP-method, since the kind of symmetry
that CMP-method exploits in a one/two-level system is not present in a
hierarchical system.

@@

Figure-8: tags for non-speculative loads:
Our invariant is that there can be only one outstanding non-speculative load,
though there can be several outstanding speculative loads concurrently.
Anything without a tag is the current non-speculative load, and anything with a
tag is a speculative load.

Reviewer-E:

Pipelining protocol-engines:
While atomic transitions in our paper have one-at-a-time semantics, the
Bluespec compiler generates hardware which executes independant transitions
concurrently in a clock cycle (as the semantics aren't violated); the compiler
excels at this optimization. Atomic transition can also be spread across
multiple clock cycles, while maintaining the semantics [4].

Beyond 500 words:
Evidence of machine-checked proofs:
We submitted our Coq source code as a supplement to our paper in the submission
website.

### 2015-01-21 Rebuttal2.txt

One of the major benefits of our approach is that it unifies the modularization
of the design and its verification task which enables verification to be done
in tandem with micro-architectural design (as noted by Reviewer-E). Further,
this allows us to require only incremental verification effort for an
incremental micro-architectural change, enabling scaling of the verification
task.

@@

Our approach vs. Model-Checking-based:

a) Deriving Invariants:
Finding invariants is fundamentally hard. The CMP-method (Reviewer-E) uses
symmetry to abstract the system, and then inserts "non-interference-lemmas"
manually to eliminate counter-examples that violate the required property.
Flow-based methodology [1] gives yet another way of manually specifying
invariants. In general, we believe that the level of complexity of the lemmas
in the two approaches are similar.  For example, for the FLASH protocol, one of
the non-interference-lemma (Lemma-5 in [2]) says that whenever a cache is in
E-state, then its data is up-to-date which is similar to our Lemma-1. The
invariants are subtle but are intuitively understood at the time of design of
each module, making the task of formalizing them during the module's
verification straightforward.

The verification of the hierarchical protocol in [3] is restricted to 2-levels
whereas our work applies to arbitrary levels of hierarchy. We are not sure if
it's a fundamental restriction on the CMP-method, since the kind of symmetry
that CMP-method exploits in a one/two-level system is not present in a
hierarchical system.

b) Case-Analysis:
Much of the requisite case-analysis is done automatically via proof-script
automation in Coq. In our experience, cases where proof-script automation fail
required an additional invariant; we believe that manually providing a similar
invariant is necessary even in a model-checking approach.

@@

Reviewer-F:

Figure-3, no tags in Figures-5,6:
Both these points are valid and we should have indicated them explicitly.

Figure-6, third-case:
The "upd" function merely determines if the "real" state of the system is as
dictated by (s,pc) or if it should be modified to take into account memory
responses. If the processor is not waiting for a memory response, then the
state is exactly (s,pc). We made an optimization in the fourth case where we
make the real state H whenever we know that the next instruction we decode is
H; we could have it as (s,pc) instead.

Figure-8: tags for non-speculative loads:
Our invariant is that there can be only one outstanding non-speculative load
(ie. loads without tags), though there can be several outstanding speculative
loads (ie. loads with tags) concurrently.

@@

Reviewer-E:

Pipelining protocol-engines:
While atomic transitions in our paper have one-at-a-time semantics, the
Bluespec compiler generates hardware which executes independent transitions
concurrently in a clock cycle (as the semantics aren't violated); the compiler
excels at this optimization. Atomic transition can also be spread across
multiple clock cycles, while maintaining the semantics [4].

@@

Evidence of machine-checked proofs:
We submitted our Coq source code as a supplement to our paper in the submission
website.

On Describing LTSes:
We include the LTS description for the benefit of readers unfamiliar with it;
the point wasn't to present novel concepts.

Beyond 500 words:
References:
[1] Parameterized Verification using Message Flows: An Industrial Experience,
John O'Leary, Murali Talupur, Mark R. Tuttle, FMCAD'09
[2] A Simple Method for Parameterized Verification of Cache Coherence Protocols,
Ching-Tsun Chou, Phanindra K. Mannava, Seungjoon Park, FMCAD'04
[3] Efficient methods for formally verifying safety properties of hierarchical
cache coherence protocol, Xiaofang Chen, Yu Yang, Ganesh Gopalakrishnan,
Ching-Tsun Chou, FMSD'10
[4] A new synthesis procedure for atomic rules containing multi-cycle function
blocks, Michal Karczmarek, Arvind, Muralidaran Vijayaraghavan, MEMOCODE'14

### 2015-01-21 Rebuttal2.txt

One of the major benefits of our approach is that it unifies the modularization
of the design and its verification task, which enables verification to be done
in tandem with micro-architectural design (as noted by Reviewer-E). Further,
this allows us to require only incremental verification effort for an
incremental micro-architectural change, enabling scaling of the verification
task.

@@

Reviewer-E:

Our approach vs. Model-Checking-based:

a) Deriving Invariants:
Finding invariants is fundamentally hard. The CMP-method (Reviewer-E) uses
symmetry to abstract the system, and then inserts "non-interference-lemmas"
manually to eliminate counter-examples that violate the required property.
Flow-based methodology [1] gives yet another way of manually specifying
invariants. In general, we believe that the level of complexity of the lemmas
between those approaches and ours are similar.  For example, for the FLASH protocol, one of
the non-interference-lemmas (Lemma-5 in [2]) says that whenever a cache is in
E-state, then its data is up-to-date, which is similar to our Lemma-1. The
invariants are subtle but are intuitively understood at the time of design of
each module, making the task of formalizing them during the module's
verification straightforward.

@@

The verification of the hierarchical protocol in [3] is restricted to 2-levels
whereas our work applies to arbitrary levels of hierarchy. We are not sure if
it's a fundamental restriction on the CMP-method, since the kind of symmetry
that CMP-method exploits in a one/two-level system is not present in a
hierarchical system.

b) Case-Analysis:
Much of the requisite case-analysis is done automatically via proof-script
automation in Coq. In our experience, cases where proof-script automation fails
required an additional invariant; we believe that manually providing a similar
invariant is necessary even in a model-checking approach.

Pipelining protocol-engines:
While atomic transitions in our paper have one-at-a-time semantics, the
Bluespec compiler generates hardware which executes independent transitions
concurrently in a clock cycle (as the semantics aren't violated); the compiler
excels at this optimization. Atomic transitions can also be spread across
multiple clock cycles, while maintaining the semantics [4].

@@

Evidence of machine-checked proofs:
We submitted our Coq source code as a supplement to our paper in the submission
website.

Figure-3, no tags in Figures-5,6:
Both these points are valid and we should have indicated them explicitly.

Figure-6, third-case:
The "upd" function merely determines if the "real" state of the system is as
dictated by (s,pc) or if it should be modified to take into account memory
responses. If the processor is not waiting for a memory response, then the
state is exactly (s,pc). We made an optimization in the fourth case where we
make the real state H whenever we know that the next instruction we decode is
H; we could have it as (s,pc) instead.

Figure-8: tags for non-speculative loads:
Our invariant is that there can be only one outstanding non-speculative load
(i.e. loads without tags), though there can be several outstanding speculative
loads (i.e. loads with tags) concurrently.

### 2015-01-29 Rebuttal2.txt

One of the major benefits of our approach is that it unifies the modularization
of the design and its verification task, which enables verification to be done
in tandem with micro-architectural design (as noted by Reviewer-E). Further,
this allows us to require only incremental verification effort for an
incremental micro-architectural change, enabling scaling of the verification
task.

Reviewer-E:

Our approach vs. Model-Checking-based:

a) Deriving Invariants:
Finding invariants is fundamentally hard. The CMP-method [1] uses symmetry to
abstract the system, and then inserts "non-interference-lemmas" manually to
eliminate counter-examples that violate the required property.  Flow-based
methodology [2] gives yet another way of manually specifying invariants. In
general, we believe that the level of complexity of the lemmas between those
approaches and ours are similar. For example, for the FLASH protocol, one of
the non-interference-lemmas (Lemma-5 in [3]) says that whenever a cache is in
E-state, then its data is up-to-date, which is similar to our Lemma-1. The
invariants are subtle but are intuitively understood at the time of design of
each module, making the task of formalizing them during the module's
verification straightforward.

The verification of the hierarchical protocol in [4] is restricted to 2-levels
whereas our work applies to arbitrary levels of hierarchy. We are not sure if
it's a fundamental restriction on the CMP-method, since the kind of symmetry
that CMP-method exploits in a one/two-level system is not present in a
hierarchical system.

@@

b) Case-Analysis:
Much of the requisite case-analysis is done automatically via proof-script
automation in Coq. In our experience, cases where proof-script automation fails
required an additional invariant; we believe that manually providing a similar
invariant is necessary even in a model-checking approach.

Pipelining protocol-engines:
While atomic transitions in our paper have one-at-a-time semantics, the
Bluespec compiler generates hardware which executes independent transitions
concurrently in a clock cycle (as the semantics aren't violated); the compiler
excels at this optimization. Atomic transitions can also be spread across
multiple clock cycles, while maintaining the semantics [5].

@@

Evidence of machine-checked proofs:
We submitted our Coq source code as a supplement to our paper in the submission
website.

Figure-3, no tags in Figures-5,6:
Both these points are valid and we should have indicated them explicitly.

Figure-6, third-case:
The "upd" function merely determines if the "real" state of the system is as
dictated by (s,pc) or if it should be modified to take into account memory
responses. If the processor is not waiting for a memory response, then the
state is exactly (s,pc). We made an optimization in the fourth case where we
make the real state H whenever we know that the next instruction we decode is
H; we could have it as (s,pc) instead.

Figure-8: tags for non-speculative loads:
Our invariant ensures there can be only one outstanding non-speculative load
(i.e. loads without tags), though there can be several outstanding speculative
loads (i.e. loads with tags) concurrently.

Beyond 500 words:
References:
[1] A simple method for parameterized verification of cache coherence protocols,
Ching-Tsun Chou, Phanindra K Mannava, Seungjoon Park, FMCAD'04
[2] Parameterized Verification using Message Flows: An Industrial Experience,
John O'Leary, Murali Talupur, Mark R. Tuttle, FMCAD'09
[3] A Simple Method for Parameterized Verification of Cache Coherence Protocols,
Ching-Tsun Chou, Phanindra K. Mannava, Seungjoon Park, FMCAD'04
[4] Efficient methods for formally verifying safety properties of hierarchical
cache coherence protocol, Xiaofang Chen, Yu Yang, Ganesh Gopalakrishnan,
Ching-Tsun Chou, FMSD'10
[5] A new synthesis procedure for atomic rules containing multi-cycle function
blocks, Michal Karczmarek, Arvind, Muralidaran Vijayaraghavan, MEMOCODE'14
